---
title: "BM_HW3"
author: "Coco"
date: "10/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

##Problem 2

#1 
The first question is to assess if their body mass index (BMI) has changed 6 years after quitting smoking. Perform an appropriate hypothesis test and interpret your findings.


```{r}
heavysmoke_df <- read_csv(file = "./data/HeavySmoke.csv") 
heavysmoke_df = janitor::clean_names(heavysmoke_df) 
t.test(heavysmoke_df$bmi_6yrs,heavysmoke_df$bmi_base,paired = T)
```

#2
The investigators suspected an overall change in weight over the years, so they decided toenroll a control group of 50-64 years of age that never smoked (data NeverSmoke.csv). Perform an appropriate test to compare the BMI changes between women that quit smoking and women who never smoked. Interpret the findings.

```{r}
neversmoke_df <- read_csv(file = "./data/NeverSmoke.csv") %>% 
  janitor::clean_names()

diff_heavy = heavysmoke_df$bmi_6yrs - heavysmoke_df$bmi_base
diff_never = neversmoke_df$bmi_6yrs - neversmoke_df$bmi_base

var.test(diff_heavy,diff_never,alternative = "two.sided")

res = t.test(diff_heavy,diff_never,var.equal = FALSE, paired = FALSE)
names(res)
```

p-value = 0.105 
not enough evidence to reject the Ho.

#3 

#4
```{r}
 power.t.test(power = .90, delta = 1.3, sd=1.75, sig.level = 0.05, alternative = c("two.sided"))
 power.t.test(power = .80, delta = 1.3, sd=1.75, sig.level = 0.05, alternative = c("two.sided"))
  power.t.test(power = .90, delta = 1.3, sd=1.75, sig.level = 0.025, alternative = c("two.sided"))
   power.t.test(power = .80, delta = 1.3, sd=1.75, sig.level = 0.025, alternative = c("two.sided"))
```

## Problem 3

```{r}
knee_df = read_csv(file = "./data/Knee.csv") %>% 
  janitor::clean_names()
summary(knee_df$below)
summary(knee_df$average)
summary(knee_df$above)

knee_df = gather(knee_df, key = type, value = value, below:above) %>% 
  filter(!is.na(value)) 
knee_df$type=as_factor(knee_df$type)

res <- aov(value~type,data = knee_df)
summary(res)
qf(0.01, df1 = 2, df2 = 22)
```

perform pairwise comparisons with the appropriate adjustments (Bonferroni, Tukey, and Dunnett – ‘below average’ as reference)
```{r}
pairwise.t.test(knee_df$value,knee_df$type,p.adjust.method = 'bonferroni')
TukeyHSD(res)
library(multcomp)
dunnetttest<-glht(res, linfct=mcp(type="Dunnett"))
summary(dunnetttest)
```

##Problem 4
```{r}

```











